---
title: "The Effects of State Minimum Wage Increases on Restaurant Employment and Prices"
subtitle: "Evidence from California and New York, 2010--2023"
author: "Anjeela Budha Magar"
date: "Econ 5253 \\ Spring 2026"
output:
  beamer_presentation:
    theme: "Madrid"
    colortheme: "default"
    fonttheme: "structurebold"
    slide_level: 2
    keep_tex: true
fontsize: 10pt
header-includes:
  - \usepackage{booktabs}
  - \usepackage{caption}
  - \captionsetup{font=footnotesize}
---

```{r setup, include=FALSE}
# ---------- knit defaults ----------
knitr::opts_chunk$set(
  echo    = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.align = "center",
  fig.width = 7, fig.height = 3.8,
  dpi = 150
)

# ---------- packages ----------
# install once on your machine, then comment out:
# install.packages(c("tidyverse", "lubridate", "lmtest", "sandwich",
#                    "kableExtra", "broom", "scales"))
suppressPackageStartupMessages({
  library(tidyverse)
  library(lubridate)
  library(lmtest)
  library(sandwich)
  library(kableExtra)
  library(broom)
  library(scales)
})

# ---------- where the raw FRED CSVs live ----------
DATA_DIR <- "."   

# ---------- analysis window ----------
YEAR_MIN   <- 2010
YEAR_MAX   <- 2023
TREAT_YEAR <- 2014


# ============================================================
# 1. helpers to read each FRED csv and aggregate to annual
# ============================================================
read_fred <- function(file, value_name) {
  read_csv(file.path(DATA_DIR, file), show_col_types = FALSE) %>%
    rename(date = 1, !!value_name := 2) %>%
    mutate(date  = as.Date(date),
           year  = year(date))
}

annual_mean <- function(df, v) {
  df %>% group_by(year) %>%
    summarise("{v}" := mean(.data[[v]], na.rm = TRUE), .groups = "drop")
}

annual_january <- function(df, v) {
  df %>% filter(month(date) == 1) %>%
    select(year, all_of(v))
}


# ============================================================
# 2. load, aggregate, tag with state, stack
# ============================================================
# minimum wage (already annual)
mw <- bind_rows(
  annual_january(read_fred("STTMINWGCA.csv", "mw"), "mw") %>% mutate(state = "CA"),
  annual_january(read_fred("STTMINWGNY.csv", "mw"), "mw") %>% mutate(state = "NY"),
  annual_january(read_fred("STTMINWGAZ.csv", "mw"), "mw") %>% mutate(state = "AZ"),
  annual_january(read_fred("STTMINWGNV.csv", "mw"), "mw") %>% mutate(state = "NV"),
  # Pennsylvania has not raised its statutory minimum since 2009;
  # the FRED file is therefore not in the data folder. Hand-set:
  tibble(year  = YEAR_MIN:YEAR_MAX, mw = 7.25, state = "PA")
)

# real GSP (quarterly -> annual mean)
gsp <- bind_rows(
  annual_mean(read_fred("CARQGSPCA.csv", "gsp"), "gsp") %>% mutate(state = "CA"),
  annual_mean(read_fred("NYRQGSP.csv",   "gsp"), "gsp") %>% mutate(state = "NY"),
  annual_mean(read_fred("AZRQGSPAZ.csv", "gsp"), "gsp") %>% mutate(state = "AZ"),
  annual_mean(read_fred("NVRQGSP.csv",   "gsp"), "gsp") %>% mutate(state = "NV"),
  annual_mean(read_fred("PARQGSP.csv",   "gsp"), "gsp") %>% mutate(state = "PA")
)

# food-services employment (monthly -> annual mean)
emp <- bind_rows(
  annual_mean(read_fred("SMU06000007072200001SA.csv", "emp"), "emp") %>% mutate(state = "CA"),
  annual_mean(read_fred("SMU36000007072200001SA.csv", "emp"), "emp") %>% mutate(state = "NY"),
  annual_mean(read_fred("SMU04000007072200001.csv",   "emp"), "emp") %>% mutate(state = "AZ"),
  annual_mean(read_fred("SMU32000007072200001SA.csv", "emp"), "emp") %>% mutate(state = "NV"),
  annual_mean(read_fred("SMU42000007072200001SA.csv", "emp"), "emp") %>% mutate(state = "PA")
)

# food-away-from-home CPI (monthly -> annual mean; AZ already annual)
cpi <- bind_rows(
  annual_mean(read_fred("CUURA422SEFV.csv", "cpi"), "cpi") %>% mutate(state = "CA"),
  annual_mean(read_fred("CUURA101SEFV.csv", "cpi"), "cpi") %>% mutate(state = "NY"),
  annual_mean(read_fred("CUURA102SEFV.csv", "cpi"), "cpi") %>% mutate(state = "PA"),
  annual_mean(read_fred("CUUSA429SEFV.csv", "cpi"), "cpi") %>% mutate(state = "AZ")
  # No Nevada CPI series available in the data folder.
)


# ============================================================
# 3. build the four DiD panels
# ============================================================
build_panel <- function(states, treat_state, include_cpi = TRUE) {
  d <- mw %>% filter(state %in% states) %>%
    inner_join(gsp %>% filter(state %in% states), by = c("state", "year")) %>%
    inner_join(emp %>% filter(state %in% states), by = c("state", "year"))
  if (include_cpi) {
    d <- d %>% inner_join(cpi %>% filter(state %in% states),
                          by = c("state", "year"))
  }
  d %>% filter(between(year, YEAR_MIN, YEAR_MAX)) %>%
    arrange(state, year) %>%
    group_by(state) %>%
    mutate(
      ln_emp     = log(emp),
      ln_gsp     = log(gsp),
      ln_emp_lag = lag(ln_emp),
      treat      = as.integer(state == treat_state),
      post       = as.integer(year >= TREAT_YEAR),
      did        = treat * post,
      ln_cpi     = if ("cpi" %in% names(.)) log(cpi) else NA_real_
    ) %>%
    ungroup()
}

panel_ca_emp <- build_panel(c("CA","AZ","NV"), "CA", include_cpi = FALSE)
panel_ny_emp <- build_panel(c("NY","PA"),      "NY", include_cpi = FALSE)
panel_ca_cpi <- build_panel(c("CA","AZ"),      "CA", include_cpi = TRUE)
panel_ny_cpi <- build_panel(c("NY","PA"),      "NY", include_cpi = TRUE)


# ============================================================
# 4. estimate DiD models
# ============================================================
# y = a + b*did + state FE + year FE + controls + e
# HC1 robust SEs; cluster-robust SEs would be preferred but with
# 2-3 states they are unstable, so HC1 is the working default.
run_did <- function(df, y, controls = character()) {
  rhs <- c("did", "factor(state)", "factor(year)", controls)
  f   <- as.formula(paste(y, "~", paste(rhs, collapse = " + ")))
  m   <- lm(f, data = df)
  ct  <- coeftest(m, vcov = vcovHC(m, type = "HC1"))
  tibble(
    beta   = unname(coef(m)["did"]),
    se     = unname(ct["did", "Std. Error"]),
    p      = unname(ct["did", "Pr(>|t|)"]),
    n      = nobs(m),
    r2     = summary(m)$r.squared,
    pct    = (exp(coef(m)["did"]) - 1) * 100
  )
}

did_results <- bind_rows(
  run_did(panel_ca_emp, "ln_emp")                            %>% mutate(spec = "CA-emp: no controls"),
  run_did(panel_ca_emp, "ln_emp", "ln_gsp")                  %>% mutate(spec = "CA-emp: + ln(GSP)"),
  run_did(panel_ca_emp, "ln_emp", c("ln_gsp","ln_emp_lag"))  %>% mutate(spec = "CA-emp: + ln(GSP) + lag"),
  run_did(panel_ny_emp, "ln_emp")                            %>% mutate(spec = "NY-emp: no controls"),
  run_did(panel_ny_emp, "ln_emp", "ln_gsp")                  %>% mutate(spec = "NY-emp: + ln(GSP)"),
  run_did(panel_ny_emp, "ln_emp", c("ln_gsp","ln_emp_lag"))  %>% mutate(spec = "NY-emp: + ln(GSP) + lag"),
  run_did(panel_ca_cpi, "ln_cpi")                            %>% mutate(spec = "CA-cpi: no controls (AZ only)"),
  run_did(panel_ca_cpi, "ln_cpi", "ln_gsp")                  %>% mutate(spec = "CA-cpi: + ln(GSP)"),
  run_did(panel_ny_cpi, "ln_cpi")                            %>% mutate(spec = "NY-cpi: no controls"),
  run_did(panel_ny_cpi, "ln_cpi", "ln_gsp")                  %>% mutate(spec = "NY-cpi: + ln(GSP)")
) %>% select(spec, beta, pct, se, p, n, r2)
```

# Introduction

## What this paper asks

\textbf{The minimum wage debate has run for thirty years. Two big questions remain.}

\bigskip

\textbf{1. Employment.} Do large state minimum wage increases reduce
restaurant jobs, as the textbook model predicts, or not, as
\textit{Card and Krueger (1994)} argued?

\bigskip

\textbf{2. Prices.} If firms don't lay workers off, do they pass the
cost on to consumers through higher menu prices?

\bigskip

\textbf{Setting.} California and New York both began phasing in
\$15-per-hour minimum wages in 2014. These are among the largest
state-level wage shocks in U.S. history --- a sharp natural
experiment.

\bigskip

\textbf{What I do.} Two parallel difference-in-differences models,
one for each treated state, on annual food-services employment and
food-away-from-home CPI, 2010--2023.

# Data

## Data sources and state pairings

\small
All series come from FRED.

```{r data-table, results='asis'}
tribble(
  ~Variable,                        ~Source,                   ~Frequency,    ~`Annualized by`,
  "State minimum wage",             "STTMINWG{CA,NY,AZ,NV}",    "Annual",      "January value",
  "Real GSP",                       "{CA,NY,AZ,NV,PA}RQGSP",    "Quarterly",   "Annual mean",
  "Food-services employment",       "SMU000007072200001SA",   "Monthly",     "Annual mean",
  "CPI: food away from home",       "CUURA{422,101,102}SEFV",   "Monthly",     "Annual mean",
  "CPI: Phoenix-Mesa (AZ proxy)",   "CUUSA429SEFV",             "Annual",      "(none)"
) %>%
  kable("latex", booktabs = TRUE, linesep = "",
        caption = "Data inputs and aggregation.") %>%
  kable_styling(font_size = 8, latex_options = "scale_down")
```

\textbf{Two state pairings:}
\begin{itemize}
  \item \textbf{California} (treated) vs \textbf{Arizona + Nevada} (control)
        --- shared sunbelt geography, tourism-heavy hospitality, similar
        seasonal patterns. AZ and NV moved their wage floors much more
        slowly than CA after 2014.
  \item \textbf{New York} (treated) vs \textbf{Pennsylvania} (control) ---
        long shared border, integrated commuter labor markets,
        similar Mid-Atlantic industrial mix. PA stayed at the federal
        floor of \$7.25 throughout.
\end{itemize}

\textbf{Caveat:} no Nevada metro food-away-from-home CPI series exists
in FRED, so the CPI specification for California uses Arizona alone
as a control.

## Minimum wage trajectories

```{r, fig.width=6, fig.height=4, out.width="0.9\\linewidth"}
mw_plot_data <- mw %>% filter(between(year, YEAR_MIN, YEAR_MAX))

ggplot(mw_plot_data, aes(year, mw, color = state, linetype = state)) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 1.6) +
  geom_vline(xintercept = TREAT_YEAR, linetype = "dashed",
             color = "gray40", linewidth = 0.4) +
  annotate("text", x = TREAT_YEAR + 0.15, y = 14.5,
           label = "Phase-in begins (2014)",
           hjust = 0, size = 2.8, color = "gray30") +
  scale_color_manual(values = c(CA = "#d62728", NY = "#1f77b4",
                                AZ = "#ff7f0e", NV = "#9467bd",
                                PA = "#2ca02c")) +
  scale_linetype_manual(values = c(CA = "solid", NY = "solid",
                                    AZ = "dashed", NV = "dashed",
                                    PA = "dashed")) +
  scale_x_continuous(breaks = seq(YEAR_MIN, YEAR_MAX, 2)) +
  scale_y_continuous(labels = dollar_format()) +
  labs(x = NULL, y = "State minimum wage",
       color = "State", linetype = "State",
       caption = "Source: FRED. Treated states (CA, NY) are solid; controls dashed.") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "right",
        plot.caption = element_text(size = 7, color = "gray40"))
```

# Methods

## Difference-in-differences specification

\textbf{Estimating equation:}

$$
\ln y_{st} \;=\; \alpha + \beta \cdot (\text{Treat}_s \times \text{Post}_t)
              + \gamma_s + \delta_t + X_{st}'\pi + \varepsilon_{st}
$$

\begin{itemize}\small
  \item $\ln y_{st}$ = log food-services employment \emph{or} log
        food-away-from-home CPI in state $s$, year $t$.
  \item $\text{Treat}_s = 1$ for CA (or NY); $\text{Post}_t = 1$ for
        years $\geq 2014$.
  \item $\gamma_s, \delta_t$ = state and year fixed effects.
  \item $X_{st}$ = log real GSP (state business cycle); for employment
        I also test specifications with the one-year lag of $\ln y$.
  \item $\beta$ = average treatment effect on the treated under
        parallel trends. Standard errors are HC1 robust.
\end{itemize}

\bigskip

\textbf{Identifying assumption:} treated and control states would have
followed parallel trajectories absent the policy. I check this with
pre-2014 growth rates next slide.

## Pre-trends check

```{r pretrends, fig.height=3.2, out.width="\\linewidth"}
pretrends <- bind_rows(
  panel_ca_emp %>% mutate(panel = "CA-emp", outcome = ln_emp),
  panel_ny_emp %>% mutate(panel = "NY-emp", outcome = ln_emp),
  panel_ca_cpi %>% mutate(panel = "CA-cpi", outcome = ln_cpi),
  panel_ny_cpi %>% mutate(panel = "NY-cpi", outcome = ln_cpi)
) %>%
  group_by(panel, state) %>%
  mutate(outcome_norm = outcome - outcome[year == YEAR_MIN]) %>%
  ungroup()

ggplot(pretrends, aes(year, outcome_norm,
                      color = state, linetype = factor(treat))) +
  geom_vline(xintercept = TREAT_YEAR, linetype = "dashed",
             color = "gray40", linewidth = 0.3) +
  geom_line(linewidth = 0.9) +
  facet_wrap(~ panel, ncol = 4, scales = "free_y") +
  scale_color_manual(values = c(CA = "#d62728", NY = "#1f77b4",
                                AZ = "#ff7f0e", NV = "#9467bd",
                                PA = "#2ca02c")) +
  scale_linetype_manual(values = c(`0` = "dashed", `1` = "solid"),
                         guide = "none") +
  scale_x_continuous(breaks = c(2010, 2014, 2018, 2022)) +
  labs(x = NULL, y = "Log change from 2010", color = NULL,
       caption = "Solid = treated state. Dashed vertical line = 2014 phase-in.") +
  theme_minimal(base_size = 9) +
  theme(legend.position = "bottom",
        plot.caption = element_text(size = 7, color = "gray40"),
        strip.text = element_text(face = "bold"))
```

\small
\textbf{Read:} CA, CA-cpi, and NY-cpi panels track their controls
closely before 2014 --- parallel-trends assumption looks credible.
The \textbf{NY-emp} panel does not: New York food-services employment
was already growing about 2.9 pp/yr faster than Pennsylvania pre-2014.
NY-emp results need to be read with that caveat.

# Findings

## Main DiD results

```{r results-table, results='asis'}
did_results %>%
  mutate(
    beta = sprintf("%+.4f", beta),
    pct  = sprintf("%+.2f%%", pct),
    se   = sprintf("%.4f",   se),
    p    = sprintf("%.3f",   p),
    r2   = sprintf("%.3f",   r2)
  ) %>%
  rename(Specification = spec, `DiD coef.` = beta, `% effect` = pct,
         `SE (HC1)` = se, `p-value` = p, `N` = n, `R²` = r2) %>%
  kable("latex", booktabs = TRUE, linesep = "",
        align = c("l","r","r","r","r","r","r"),
        caption = "DiD coefficient on (Treat $\\times$ Post).") %>%
  kable_styling(font_size = 7, latex_options = c("scale_down","HOLD_position")) %>%
  row_spec(0, bold = TRUE) %>%
  pack_rows("Employment outcomes (log food-services employment)", 1, 6,
            latex_gap_space = "0.4em") %>%
  pack_rows("Price outcomes (log CPI food-away-from-home)", 7, 10,
            latex_gap_space = "0.4em")
```

\small
\textbf{Three patterns to highlight:}
\begin{itemize}\small
  \item \textbf{CA employment $\downarrow$ 2--6\%} relative to AZ+NV ---
        the textbook prediction shows up here.
  \item \textbf{CA prices $\uparrow$ 5--7\%} relative to AZ ---
        substantial pass-through of wage costs to menu prices.
  \item \textbf{NY results are mixed} --- positive employment
        coefficient but pre-trends fail the test; price effect is
        small and shrinks once we control for state GSP.
\end{itemize}

## Visual evidence: California employment

```{r, fig.width=6, fig.height=4, out.width="0.9\\linewidth"}
plot_emp <- panel_ca_emp %>%
  group_by(state) %>%
  mutate(emp_norm = emp / emp[year == YEAR_MIN] * 100) %>%
  ungroup()

ggplot(plot_emp, aes(year, emp_norm, color = state, linetype = state)) +
  geom_vline(xintercept = TREAT_YEAR, linetype = "dashed",
             color = "gray40", linewidth = 0.4) +
  geom_line(linewidth = 1.1) +
  geom_point(size = 1.6) +
  scale_color_manual(values = c(CA = "#d62728",
                                AZ = "#ff7f0e", NV = "#9467bd")) +
  scale_linetype_manual(values = c(CA = "solid",
                                    AZ = "dashed", NV = "dashed")) +
  scale_x_continuous(breaks = seq(YEAR_MIN, YEAR_MAX, 2)) +
  labs(x = NULL, y = "Food-services employment (2010 = 100)",
       color = "State", linetype = "State",
       caption = "California (solid) tracks the AZ-NV average closely until 2014, then diverges downward.") +
  theme_minimal(base_size = 10) +
  theme(legend.position = "right",
        plot.caption = element_text(size = 7, color = "gray40"))
```

# Conclusion

## What did we learn?

\textbf{Three takeaways for California:}
\begin{itemize}
  \item Modest \textbf{disemployment} ($-2$ to $-6\%$) in food
        services, in line with the recent literature on large wage
        shocks (\textit{Jardim et al.\ 2022, Seattle}).
  \item Substantial \textbf{price pass-through} ($+5$ to $+7\%$) ---
        consistent with \textit{Aaronson (2001)} and
        \textit{Harasztosi \& Lindner (2019)}: firms absorb wage shocks
        mostly through prices, partly through employment.
  \item Net effect on workers is ambiguous: those who keep their
        jobs earn more; those displaced lose; consumers pay more.
\end{itemize}

\bigskip

\textbf{Why it matters:}
\begin{itemize}
  \item Federal and state minimum wage debates (\$15+ proposals
        nationwide) need credible estimates of \emph{both}
        employment and price effects.
  \item Single-outcome studies miss half the story --- a paper that
        looks only at jobs would conclude ``not much happened''; one
        that looks only at prices would conclude ``consumers paid for
        it.'' The truth is both.
\end{itemize}

\bigskip

\textbf{Honest caveats:} small number of states (low power),
state-aggregate data masks within-state variation (NYC vs upstate),
NY pre-trends fail the parallel test. County-level QCEW data and
synthetic-control methods would strengthen the design.
